 ---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->

```{r, echo = FALSE}
version <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
version <- gsub('-', '.', version)                          
```
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![packageversion](https://img.shields.io/badge/Package%20version-`r version`-orange.svg?style=flat-square)](commits/master)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)](/commits/master)

<!-- badges: end -->


## Utility

**soiltestr** is built to help collect and analyze laboratory soil test data. I wrote the functions for my own use, but other soil scientists may find them of value. The code streamlines commonly performed tests: particle size analysis, Atterberg limits, Proctor maximum density, and particle density (specific gravity).

This package is currently under heavy development and likely will be rather opaque to new users. However, I plan to streamline the documentation and release it to CRAN by July 2021.

## Installation

A binary version of `soiltestr` is not yet available on CRAN. Install the development version with the following command:

``` r
remotes::install_github("evanmascitti/soiltestr")
```

<!-- The best way to learn about this package is through the vignettes  -->
<!-- un-comment this line once the vignettes are added  -->

## Functions 

Below are one-line overviews of each function's purpose:

### Data analysis 

`add_w()`: Calculate gravimetric water content and add to an existing data frame

`add_physical_properties()` Compute various metrics from raw Proctor data

`proctor_prep()`: Specimen preparation for Proctor compaction tests

`proctor_fit()`: Fit a compaction curve to compute w<sub>opt</sub> and &rho;<sub>max</sub> 

`d_max()` Fit a compaction curve and compute the maximum density only.  

`w_opt()` Fit a compaction curve and compute the optimum water content only. 

[`compute_LL()`](R/compute_LL.R) calculate the liquid limit from water contents and blow counts

[`LL_batch_analysis()`](R/LL_batch_analysis.R) calculate liquid limit of all specimens tested in a batch

[`PL_batch_analysis()`](R/PL_batch_analysis.R) calculate plastic limit of all specimens tested in a batch

[`AL_batch_analysis()`](R/AL_batch_analysis.R) calculate adhesion limit of all specimens tested in a batch

[`pipette_analysis()`](R/pipette_analysis.R) analyze pipette and sieve data for one or more samples


### Visualization 

`ggproctor()` plotting compaction curves with **ggplot2**

`ggflowcurve()` plotting liquid limit data with **ggplot2**

`casagrande_chart()` display Atterberg limits on the traditional LL vs PI scatterplot, with U-line and A-line plotted

[`ggpsd()`](R/ggpsd.R) plot one or more particle size distributions

[`ggpsd_single_sample()`](ggpsd_single_sample.R.R) plot particle size distributions for a single specimen

[`ggphase_diagram()`](R/gg_phase_diargam.R) plot solid, liquid, and gaseous phases of a sand-clay mix


### Data managmement and reference sheets

`mix_calcs()`: Precise calculations for sand-clay soil mixtures

#### Populate skeleton .csv files for data collection (function names self-explanatory)

`generate_proctor_datasheet()` 

[`pipette_datasheets()`](R/pipette_datasheets.R) creates .csv data sheets for the different types of data required in this analysis

[`att_lims_datasheets()`](R/att_lims_datasheets.R) creates .csv data sheets for liquid, adhesion, and plastic limit tests


## Standard method references

The standardized soil tests which may be deployed using this package include: 

#### Particle size analysis
* Hydrometer method: [ASTM D7928-17](https://www.astm.org/Standards/D7928.htm)
* Pipette method: [SSSA Methods of Soil Analysis, Part 4](https://www.wiley.com/en-us/Methods+of+Soil+Analysis%2C+Part+4%3A+Physical+Methods-p-9780891188933)

#### Proctor compaction
* Standard effort: [ASTM D698-12e2](https://www.astm.org/Standards/D698.htm)
* Modified effort: [ASTM D1557-12e1](https://www.astm.org/Standards/D1557)

#### Atterberg limits 
* Canonical thread-rolling/Casagrande cup: [ASTM D4318-17e1](https://www.astm.org/Standards/D4318)
